Draft_Application_Stochastic_Volatility

### 2021-07-20 Real Dataset SV.Rmd
commit

When focusing the attention on specific filters, first we can discuss the effect of resampling. Indeed, in the plot referring to the Sequential Importance Sampling procedure, the sequence of filtered volatilities is not able to reproduce the peaks as in with other filters, but also is characterized by collapsing confidence intervals. This is because ::::::: \
Then, if we instead look at the filtered volatilities obtained with resampling, we can appreciate how the filtered sequences for the Basic,\footnote{By this name, we refer to a particle filter that features no adaptive resampling; rather, resampling is conducted at each step of the algorithm.} Bootstrap, Auxiliary and Optimal Kernel Guided particle filters give rather analogous results, with no differences appearing at a first visual inspection.\footnote{Except for the Auxiliary Particle filtered sequence of volatilities, showing a slightly higher peak corresponding to the greatest volatility fluctuation in 2020, if compared to other filters in this subset. This is still valid for all three indices.} \
However, the sequence of filtered volatilities obtained by applying the Liu and West filter is clearly different from the others. Indeed, it shows a "more detailed" sequence of states (in the sense of the sequence of filtered volatilities being less smooth) and also a higher peak for the wide fluctuation corresponding to the 2020 Covid Crisis outbreak. Therefore, we may say that, compared to other filtering techniques, it offers a better representation of the data. This is because ... \textit{Potremmo parlare di maggiore sensibilità alle oscillazioni dei prezzi, o simili}

@@

...

>>>>

When focusing the attention on specific filters, first we can discuss the effect of resampling. Indeed, in the plot referring to the Sequential Importance Sampling procedure, the sequence of filtered volatilities is not able to reproduce the peaks as in with other filters, but also is characterized by collapsing confidence intervals. This is because ::::::: \
Then, if we instead look at the filtered volatilities obtained with resampling, we can appreciate how the filtered sequences for the Basic,\footnote{By this name, we refer to a particle filter that features no adaptive resampling; rather, resampling is conducted at each step of the algorithm.} Bootstrap, Auxiliary and Optimal Kernel Guided particle filters give rather analogous results, with no differences appearing at a first visual inspection.\footnote{Except for the Auxiliary Particle filtered sequence of volatilities, showing a slightly higher peak corresponding to the greatest volatility fluctuation in 2020, if compared to other filters in this subset. This is still valid for all three indices.} \
However, the sequence of filtered volatilities obtained by applying the Liu and West filter is clearly different from the others. Indeed, it shows a "more detailed" sequence of states (in the sense of the sequence of filtered volatilities being less smooth) and also a higher peak for the wide fluctuation corresponding to the 2020 Covid Crisis outbreak. Therefore, we may expect that, compared to other filtering techniques, it can provide a better representation of the data. This is because ... \textit{Potremmo parlare di maggiore sensibilità alle oscillazioni dei prezzi, o simili}

@@

In section XX we compared such distributions already, pointing out three main behaviours, namely that of the SIS algorithm, that of the Liu and West particle filter and that of the remaining filters; however, we did not have metrics that could help us clearly point at one behaviour as the closest to reality. We will now use a realised volatility measure as one such metric.

\subsection{Comparing the filters}

In order to evaluate the different filters according to the evaluation tool represented by the realized volatility, we start once more from the visual inspection of the plots. For any of the three indices, below graphs are shown including both the filtered volatility obtained from the filtered states (for any of the Sequential MC methods described in the theoretical section) and the corresponding rescaled realised volatility series.\footnote{For simplicity,troughout the rest of this section we will omit "rescaled" when refering to the rescaled realised volatility.} With respect to the visual analysis we conducted in section \textit{fill with the number of the section}, we can now use such a proxy as to determine the relative performance of the competing filters.\

Once more, most algorithms used (aside from SIS and Liu-West) seem to provide indistinguishable results and, generally speaking, to underestimate the volatility (as indicated by the realised volatility proxy). The SIS algorithm appears completely inadequate to describe the "true" volatility: as we argued in section XX, the degeneracy of its weights on the particles prevents it from delivering a good enough approximation of the target distribution (\textit{check whether this is the correct intuition}. On top of that, it does not seem the case that the SIS algorithm, by falling short of the SV target distribution, ends up fitting the proxy better.\footnote{As we will now see, this seems instead to be the case for our preferred algorithms, namely those adopting adaptive resampling rules: a better approximation of the SV target filtering distribution yields slightly higher RMSEs and MAEs than those resulting from less accurate MC approximations.}\
On the other hand, the Liu and West filter offers the best performance among all filters, at least visually. This suggests that, since the Liu and West particle filter is built on a different underlying model than the other filters considered, this model may offer a better representation of the data analyzed. Once more, these considerations hold for all of the three indices.\

GRAFICI

The above intuitions are confirmed by looking at Table XX, where the RMSE and MAE are shown for each filter (run with a number of particles equal to 10000), considering the realised volatility as benchmark. Indeed, for any index, most of the filters give rather similar results in terms of RMSE and MAE. For any index and for any measure, the worst performing is the SIS algorithm, while the best performing is the Liu and West particle filter. (Metterei qua l'intuizione del Liu West)\

```{r include=FALSE}
Errorcomp3<-matrix(NA,ncol=7,nrow=6)
colnames(Errorcomp3)<-c("N","BPF", "GPFOPT", "APF", "BAPF","LW", "SIS")
rownames(Errorcomp3)<-c("RMSE SP500","MAE SP500","RMSE STOXX","MAE STOXX","RMSE DJI","MAE DJI")
Errorcomp3[,1]<-c(10000,10000,10000,10000,10000,10000)
Errorcomp3[,2]<-c(0.508642137861919,0.353767519133916,0.463584874517798,0.301898325469172,0.507624385776376,0.337543646258033)
Errorcomp3[,3]<-c(0.509256032254369,0.35314674168209,0.463446053320389,0.304178601158929,0.506222076244991,0.338268970683459)
Errorcomp3[,4]<-c(0.509943715852212,0.351738603428736,0.475708272485282,0.30422270387141,0.509343139732461,0.337449244130176)
Errorcomp3[,5]<-c(0.510334461444389,0.35305864294958,0.46783471248187,0.304213895457104,0.505360139532104,0.3361502165818)
Errorcomp3[,6]<-c(0.428716693039206,0.291557534456417,0.344065170120452,0.237159281665533,0.440832171188778,0.280952286978564)
Errorcomp3[,7]<-c(0.602883459334495,0.431972177956328,0.652533564914959,0.440767074642285,0.604672122750974,0.415631855223678)

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
library(kableExtra)
kbl(Errorcomp3, align="c", longtable = T, booktabs = T, caption="RMSE and MAE, realised volatility as benchmark",digits=4)
```

Finally, we can see from Table XX the RMSE and MAE obtained with a number of particles equal to 50000 for the SP500 index (\textit{we will consider whether it is worth it to include other indexes: for the moment, I would stick to the SP500, as when considering N=10000 we are already in the convergence phase}). In this case, although the number of particles increases, the RMSE and MAE increase (although by very small amounts). This is evidence supporting the idea that the model we have selected underlying for the first two filters is inappropriate for describing this specific data. Indeed, as widely discussed in SECTION XX, it is obsolete and surpassed by more advanced and precise models.

```{r include=FALSE}
Errorcomp4<-matrix(NA,ncol=4,nrow=2)
colnames(Errorcomp4)<-c("N","BPF", "APF","LWF")
rownames(Errorcomp4)<-c("RMSE","MAE")
Errorcomp4[,1]<-c(50000,50000)
Errorcomp4[,2]<-c(0.511700619467058,0.354186384429001)
Errorcomp4[,3]<-c(0.514888164962505,0.355336840861771)
Errorcomp4[,4]<-c(0.4242305,0.2906932)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
library(kableExtra)
kbl(Errorcomp4, align="c", longtable = T, booktabs = T, caption="RMSE and MAE, realised volatility as benchmark (SP500)",digits=5)
```
